---
title: "L12_Analysis"
author: "Weinberg"
date: "2022-11-07"
output: html_document
---
•	Graphs and the results of statistical tests addressing whether or not there is a significant difference in classical conditioning response across life history stage, sex, and individual identity.

Test for effect of the following on learning ability (MLR response to CS+ in retention test):
•	Age
•	Sex
•	Individual variation


#Box 0. Import the data and libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, ggplot2, readxl)

dat<-read_excel("L10_LearningMemory_FA23.xlsx")
colnames(dat)<-c("group", "animal", "sex", "instar", "length", "CS_plus", "CS_minus", "control")
dat<-na.omit(dat)
dat
```

#Box 1. View the data and summary statistics
In order to classify an individual as having correctly learned the conditioned stimulus, it must:
-Exhibit the MLR when exposed to the CS+ odor
-Not exhibit MLR when exposed to the CS- or control odor

Print the percentage of cockroaches who learned the stimulus
-By sex 
-By instar (age)

```{r}
dat$result<-0
for(i in 1:nrow(dat)){
  if(dat[i,6]==1 & dat[i,7]==0 & dat[i,8]==0){dat$result[i]<-1}
  else{dat$result[i]<-0}
}
perc<-function(x){sum(x)/length(x)*100}

sumsex<-dat %>% group_by(sex) %>% summarize(percent.learned=perc(result))
suminstar<- dat %>% group_by(instar) %>% summarize(percent.learned=perc(result))
sumsex
suminstar

```

#Box 2a. Plot the data by sex
```{r}
plot_sex<-ggplot(data = sumsex, aes(x=sex, y=percent.learned, fill=sex))+
  geom_col()+labs(title="Percentage of Cockroaches Correctly Learning Conditioned Stimulus by Sex")
plot_sex
```

#Box 2b. Plot the data by instar
Replace "my plot title" in the first line below with the title you would like to use for this plot.
```{r}
dat$instar<-as.factor(dat$instar)
plot_instar<-ggplot(data = suminstar, aes(x=instar, y=percent.learned, fill=instar))+
  geom_col()+labs(title="Percentage of Cockroaches Correctly Learning Conditioned Stimulus by Age")

plot_instar
```
#Box 2c. Plot individual learning responses 

In this plot, each individual has its own value on the x-axis. A value of "1" means that the individual learned the behavior, while a value of "0" means that it did not.

```{r}
plot_individual<-ggplot(data = dat %>% mutate(ID=row_number()), aes(x=ID, y=result))+
  geom_point(aes(shape=instar, color=sex))+labs(title="Individual Learning Responses for Madagascar Hissing Cockroaches")

plot_individual
```

#Box 3. Run statistical tests (ANOVA) to investigate whether differences in learning ability vary significantly with:
-age
-sex
-combination of age and sex

Based on the p-values, do you see a signficant difference in the percentage of cockraoches correctly learning the stimulus based on age, sex, or both?
```{r}
dat$sex<-as.factor(dat$sex)
sex.aov<-aov(result~sex, data=dat)

instar.aov<-aov(result~instar, data=dat)

intx.aov<-aov(result~sex*instar, data=dat)


anovas<-data.frame(effects=c("age", "sex", "age*sex"), p_value=c(summary(instar.aov)[[1]]$`Pr(>F)`[1],
                                                                   summary(sex.aov)[[1]]$`Pr(>F)`[1],
                                                                   summary(intx.aov)[[1]]$`Pr(>F)`[1]))
anovas
```
