---
title: "L2+3_pollinators"
author: "Weinberg"
date: "2022-08-13"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

#Box 0. Run this whole chunk to load the libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(tidyverse, ggplot2, readxl, RColorBrewer)
```

# Box 1. Import the data

1. Replace "my name" (first line) with your name, **exactly as you typed it on the tab for the lab 2 spreadsheet** (see slides from this week). Only change the text in quotation marks. Make sure you enter your name inside the quotation marks.

2. Replace "my group" (second line) with your group name, **exactly as you typed it on the tab for the lab 3 spreadsheet** (see slides from this week). Only change the text in quotation marks. Make sure you enter your name inside the quotation marks.

3. After you have changed those lines, run the cell to make sure the data look correct. Click the boxes that appear below the cells to see each of the two data tables that you imported
```{r import data}
name<-"my name"
team<-"my group"

l2<-read_excel('L02 Real Flower Data.xlsx', sheet=name)
l3<-read_excel('L03 Artificial Flower Data.xlsx', sheet=team)
l2$Behavior<-tolower(l2$Behavior)
l2[is.na(l2)]<-0
l3[is.na(l3)]<-0
l3$Behavior<-tolower(l3$Behavior)
l2
l3
```

#Box 2. Organize the data into one table. The output should be a data table with the first 20 rows holding the data for your week 1 real flower observations and the next 32 rows holding the data for your artificial flowers 

```{r}
dat<-rbind(l2, l3[,-2])
dat <-cbind(Type=c(rep("Real", nrow(l2)), rep("Artificial", nrow(l3))), dat)
dat
```

#Box 3. 
Look at some of the summary statistics:
First table: average number of observations for each behavior on each flower type and standard error
Second table: average number of behaviors recorded for each insect taxon on each flower type and standard error

```{r message=FALSE}
se=function(x){sd(x)/sqrt(length(x))}

SumStats_behavior<-dat %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Type, Behavior) %>% summarise(Average=mean(Observations), SE=se(Observations))

SumStats_behavior

SumStats_taxa<-dat %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Type, Insect) %>% summarise(Average=mean(Observations), SE=se(Observations))

SumStats_taxa

SumStats_artificial<-l3 %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Flower, Insect) %>% summarise(Average=mean(Observations), SE=se(Observations))
SumStats_artificial
```

#Box 4. Start by creating plots showing the incidences of each behavior on each of your real and artificial flowers. The first plot will show all the behaviors recorded on the real flower, the second plot will show the behaviors recorded on the artificial flower, and the third plot will show the averages over all observations to allow for comparison between the real and artificial flowers. In these plots, all insect taxa are combined. We will look at behaviors by insect type in the next box.

Note that the number 1-4 after each flower type in the artificial flower plot represent the four different observation periods.

```{r message=FALSE}
taxa<-colnames(dat[4:ncol(dat)])
beh_lower<-SumStats_behavior$Average-SumStats_behavior$SE
beh_upper<-SumStats_behavior$Average+SumStats_behavior$SE


real_flowers<-ggplot(data=dat[1:nrow(l2),] %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Flower), aes(x=Flower, y=Observations, fill=Behavior))+
  geom_col(position="dodge")+
  ggtitle("Observed Behaviors on Real Flowers")

real_flowers

artificial<-ggplot(data=dat[(nrow(l2)+1):nrow(dat),] %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Flower), aes(x=Flower, y=Observations, fill=Behavior))+
  geom_col(position="dodge") + theme(axis.text.x=element_text(angle=45, hjust=1), axis.title.x = element_text(vjust=0.5)) + ggtitle("Observed Behaviors on Artificial Flowers")

artificial


combined<-ggplot(data=dat %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Type, Behavior) %>% summarise(Average=mean(Observations), SE=se(Observations)), aes(x=Type, y=Average, fill=Behavior))+
  geom_col(position="dodge")+ ylim(0,4)+
  labs(title="Average Number of Observed Insect Behaviors on Real and Artificial Flowers", x="Flower Type", y="Average Number of Observations") + geom_errorbar(ymin=beh_lower, ymax=beh_upper, position = "dodge")

combined
```

#Box 5. Now we can plot the insects of each type that visited each flower. To keep this plot from getting too busy, we will just sum up the total number of observations across all the behaviors ("Observations" on the y-axis now shows the sum of all visits, landings, inspections, and feedings).

The first plot shows the total number of observations of each type of insect on real flowers. The second plot shows the same thing but on the artificial flowers. The third plot shows the average number of interactions each type of insect had with the real or artificial flowers. 

```{r, warning=FALSE, message=FALSE}
tax_upper<-SumStats_taxa$Average-SumStats_taxa$SE
tax_lower<-SumStats_taxa$Average+SumStats_taxa$SE

real_flowers_type<-ggplot(data=dat[1:nrow(l2),] %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Insect, Flower) %>% summarise(Total=sum(Observations)), aes(x=Flower, y=Total, fill=Insect))+
  geom_col(position="dodge")+
  ggtitle("Types of Insects Observed on Real Flowers") + scale_fill_brewer(palette="Set1")


real_flowers_type

#check dat here
artificial_type<-ggplot(data=dat[(nrow(l2)+1):nrow(dat),] %>% pivot_longer(taxa, names_to="Insect", values_to="Observations") %>% group_by(Insect, Flower) %>%  summarise(Total=sum(Observations)), aes(x=Flower, y=Total))+
  geom_col(position="dodge", aes(fill=Insect)) + theme(axis.text.x=element_text(angle=45, hjust=1), axis.title.x = element_text(vjust=0.5)) + ggtitle("Types of Insects Observed on Artificial Flowers") + scale_fill_brewer(palette="Set1")

artificial_type

combined_type<-ggplot(data=dat %>% pivot_longer(colnames(dat[4:ncol(dat)]), names_to="Insect", values_to="Observations") %>% group_by(Type, Insect) %>% summarise(Average=mean(Observations), SE=se(Observations)), aes(x=Type, y=Average, fill=Insect))+
  geom_col(position="dodge")+ ylim(0,6)+
  labs(title="Average Number of Observed Insects on Real and Artificial Flowers", x="Flower Type", y="Average Number of Observations") + geom_errorbar(ymin=tax_lower, ymax=tax_upper, position = "dodge") + scale_fill_brewer(palette="Set1")

combined_type
```

Box 6.
For the final plot, we'll look at just the two flowers you constructed side-by-side to see if the two designs successfully attracted different types of insects. 
```{r}
art_lower<-SumStats_artificial$Average-SumStats_artificial$SE
art_upper<-SumStats_artificial$Average+SumStats_artificial$SE
ymax<-max(art_upper)+0.5


broad_v_target<-ggplot(data=SumStats_artificial, aes(x=Flower, y=Average, fill=Insect))+
  geom_col(position="dodge") + ylim(0,ymax) + geom_errorbar(ymin=art_lower, ymax=art_upper, position="dodge") + labs(title="Average Number of Interactions by Each Insect Type on Artificial Flowers", y="Average number of Interactions", x="Artificial Flower Design") + scale_fill_brewer(palette="Set1")

broad_v_target
```


###Part 2: Statistics###

Box 7. We calculated descriptive statistics for three different groupings in Box 2 (Behaviors by flower type, taxa by flower type, and taxa on broad range or targeted artificial flowers). Those gave us a general idea of how different insects interacted with different flowers, but now we are going to run statistical tests across each of those comparisons to answer the following questions:

1.Is there a difference in the types of insect behaviors observed on real and artificial flowers?
(Landing, inspecting, feeding, visit only)

2. Is there a difference in the types of insects observed on real and artificial flowers?
(Bees and wasps, butterflies and moths, flies, and beetles)

3. Is there a difference in the types of insects that interact with the broad range and targeted artificial flower?


This first box outputs a table of t-test results to answer question #1
```{r}
dat1<-dat %>% select(Type, Behavior, all_of(taxa)) %>% group_by(Type, Behavior) %>% mutate(total=rowSums(across(c(taxa))))
#


visit<-dat1 %>% filter(Behavior=="visit")
landing<-dat1 %>% filter(Behavior=="landing")
inspecting<-dat1 %>% filter(Behavior=="inspecting")
feeding<-dat1 %>% filter(Behavior=="feeding")
                                                                                      
t_ra_visit<-t.test(total~Type, data=visit)
vt<-c("visit", t_ra_visit$estimate[[1]], t_ra_visit$estimate[[2]], t_ra_visit$p.value)

t_ra_landing<-t.test(total~Type, data=landing)
lt<-c("landing", t_ra_landing$estimate[[1]], t_ra_landing$estimate[[2]], t_ra_landing$p.value)

t_ra_inspecting<-t.test(total~Type, data=inspecting)
it<-c("inspecting", t_ra_inspecting$estimate[[1]], t_ra_inspecting$estimate[[2]], t_ra_inspecting$p.value)

t_ra_feeding<-t.test(total~Type, data=feeding)
ft<-c("feeding", t_ra_feeding$estimate[[1]], t_ra_feeding$estimate[[2]], t_ra_feeding$p.value)

beh_t<-data.frame(rbind(vt, lt, it, ft))
colnames(beh_t)<-c("Behavior", "Mean_Artificial", "Mean_Real", "P-value")
rownames(beh_t)<-NULL
beh_t
```
This box outputs a table of t-test results to answer question #2: 
Is there a difference in the types of insects observed on real and artificial flowers?

```{r}
dat2<-dat %>% pivot_longer(cols=all_of(taxa), names_to = "Insect", values_to="Total") %>% group_by(Type)

bee<-dat2 %>% filter(Insect=="bee_wasp")
butterfly<-dat2 %>% filter(Insect=="butterfly_moth")
fly<-dat2 %>% filter(Insect=="fly")
beetle<-dat2 %>% filter(Insect=="beetle")


t_ra_beewasp<-t.test(Total~Type, data=bee)
tb<-c("bee_wasp", t_ra_beewasp$estimate[[1]], t_ra_beewasp$estimate[[2]], t_ra_beewasp$p.value)
  

t_ra_butterfly<-t.test(Total~Type, data=butterfly)
tm<-c("butterfly_moth", t_ra_butterfly$estimate[[1]], t_ra_butterfly$estimate[[2]], t_ra_butterfly$p.value)

t_ra_fly<-t.test(Total~Type, data=fly)
tf<-tm<-c("fly", t_ra_fly$estimate[[1]], t_ra_fly$estimate[[2]], t_ra_fly$p.value)

t_ra_beetle<-t.test(Total~Type, data=beetle)
tt<-c("beetle", t_ra_beetle$estimate[[1]], t_ra_beetle$estimate[[2]], t_ra_beetle$p.value)

taxa_t<-data.frame(rbind(tb, tm, tf, tt))
colnames(taxa_t)<-colnames(beh_t)<-c("Insect", "Mean_Artificial", "Mean_Real", "P-value")
rownames(taxa_t)<-NULL

taxa_t
```
This box outputs a table of t-test results for question #3: Is there a difference in the types of insects that interact with the broad range and targeted artificial flower?

```{r}
dat3 <- l3 %>% pivot_longer(cols=all_of(taxa), names_to="Insect", values_to="Total") %>% group_by(Flower)

bee_a<-dat3 %>% filter(Insect=="bee_wasp")
butterfly_a<-dat3 %>% filter(Insect=="butterfly_moth")
fly_a<-dat3 %>% filter(Insect=="fly")
beetle_a<-dat3 %>% filter(Insect=="beetle")


t_bee<-t.test(Total~Flower, data=bee_a)
tbt<-c("bee_wasp",  t_bee$estimate[[1]], t_bee$estimate[[2]], t_bee$p.value)

t_butterfly<-t.test(Total~Flower, data=butterfly_a)
tmt<-c("butterfly_moth",  t_butterfly$estimate[[1]], t_butterfly$estimate[[2]], t_butterfly$p.value)

t_fly<-t.test(Total~Flower, data=fly_a)
tft<-c("fly",  t_fly$estimate[[1]], t_fly$estimate[[2]], t_fly$p.value)

t_beetle<-t.test(Total~Flower, data=bee_a)
tet<-c("beetle",  t_beetle$estimate[[1]], t_beetle$estimate[[2]], t_beetle$p.value)

taxa_tb<-data.frame(rbind(tbt, tmt, tft, tet))
colnames(taxa_tb)<-c("Insect", "Mean_Artificial", "Mean_Real", "P-value")
rownames(taxa_tb)<-NULL

taxa_tb

```



